Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management

### 2012-09-19 chi2012.tex
minor

While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
In this paper, we present an initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
These observations led to our design of DataPalette, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  
Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.

@@

\subsubsection{Study and task design}
To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).  Each task was paired with one of two datasets, depending on condition, which were counterbalanced in a full-factorial design to eliminate potential ordering, task, and dataset biases.\looseness=-1

@@

In the control conditions, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from. Participants were told they were free to use the web sites and spreadsheets however they wished.

@@

\subsubsection{Efficiency: Time taken}
Participants spent an average of 8.9 minutes performing the tasks. Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$); however, a 2-way repeated measures ANOVA test of time taken by interface condition, blocked by participant ID, revealed that interface condition did not have a significant effect on task completion time.  This analysis also showed no significant effect of participant ID on task completion.  Comparing tasks, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), but this difference was not significant.

\subsubsection{Thoroughness: Factors weighed in final choice}
The second metric was the number of factors the participant considered in making their final choice. To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.  A 2-way ANOVA test of the effect of interface and participant ID on the number of factors mentioned revealed a significant effect; post-hoc analysis with Tukey-Scheff\'{e} adjustments indicated that participants in the DP condition mentioned significantly more factors ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$. A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning some participants mentioned significantly more factors than others ($F(19,19)=4.53; p < 0.001$).

\subsubsection{Diversity: Data sources consulted}
The third metric we used to gauge the task performance is the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data. In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$) in the control condition, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p < 0.15$).

\subsubsection{Effort: External cognition}
Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much information participants had to keep track of during the study.  We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in a 2-way ANOVA blocked by participant ID ($F(1,19)= 8.22; p\approx0.10$).

\begin{figure}[tbp]
\begin{center}
\includegraphics[width=8.5cm]{img/results}
\caption{\emph{Choice picks per participant trial} - Histogram of number of times each restaurant/university was chosen in DP interface (A condition) in blue/left bars, compared to the Excel/website (B condition) in red/to the right.}
\label{fig:chosen_results}
\end{center}
\end{figure}

\subsubsection{Candidate selection}
Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.  However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.  As can be seen there is considerable variability and no discernible difference in variability between conditions.

@@

We found that the use of DP strongly influenced people to keep all candidates around, while participants in the B condition eliminated choices early.  An ANOVA test demonstrated a significant effect between condition and number of candidates maintained; a post-hoc analysis with Tukey-Scheff\'{e} adjustments confirmed that the number of candidates used in the final choice for participants in the DP condition was greater ($M=5.85;SD=0.45$) than the control condition ($M=4.95; SD=2.26$) $F(1,19)=5.323; p<0.05$.

@@

As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused ``ad-hoc'' approaches to be overlooked.  Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones. Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.  Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through easy to use, and intuitive interactive approaches for end-users.

>>>>

While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies examining the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
These observations led to our design of DataPalette, an interface that introduced simple drag-and-drop and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.  
Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition during multidimensional, subjective-choice decisions.

@@

\subsubsection{Study and task design}
To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).  Each task was paired with one of two datasets, depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.\looseness=-1

@@

In the control condition (B), participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from. Participants were told they were free to use the web sites and spreadsheets however they wished.

@@

\subsubsection{Efficiency: Time taken}
Participants spent an average of 8.9 minutes performing the tasks. Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).  An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.  Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.

\subsubsection{Thoroughness: Factors weighed in final choice}
The second metric was the number of factors the participant considered in making their final choice. To determine this, we asked participants to explain precisely why they made their choice, and matched their justifications to properties in the data.  This analysis revealed the interface used made a significant impact on the number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.

\subsubsection{Diversity: Data sources consulted}
The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial. In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).

\subsubsection{Effort: External cognition}
Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.  We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8.5cm]{img/results}
\caption{\emph{Choice picks per participant trial} - Histogram of number of times each restaurant/university was chosen in DP interface (A condition) in blue/left bars, compared to the Excel/website (B condition) in red/to the right.}
\label{fig:chosen_results}
\end{center}
\end{figure}

\subsubsection{Candidate selection}

Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.  However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.  As can be seen there is considerable variability and no discernible difference in variability between conditions.

@@

We found that the use of DP strongly influenced people to keep all candidates around, while participants in the B condition eliminated choices early.  An ANOVA test demonstrated a significant effect between condition and number of candidates maintained ($F(1,19)=5.323; p<0.05$); a post-hoc Tukey HSD test confirmed that the number of candidates used in the final choice for participants in the A condition ($M=5.85;SD=0.45$) differed significantly from the B condition ($M=4.95; SD=2.26$).

@@

As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused ``ad-hoc'' approaches to be overlooked.  Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones. Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.  Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through simple, fast interactive approaches that are received as intuitive by end-users.
